---
title: "ProjectCourse6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Dit is een R Markdown document voor het project van course 6 gemaakt door groep 1. blablabla

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:


```{r include = FALSE}
library(edgeR)
require(limma)
```
### Huidige werkmap instellen naar map waar script in staat. 
```{r}
#setwd(dirname(rstudioapi::getSourceEditorContext()$path)) 
fDir <-  file.choose()
```
### read counts
```{r}

inhoud = read.table("./RNAseq-bestanden/RNA-Seq-counts.txt", header=TRUE)
#row.names(inhoud) = inhoud[,"ID"]
annotatie = read.table(annotatiebestand, header=TRUE, sep="\t"
                  , quote="")
names(annotatie)[names(annotatie) == "name"] = "ID"
data = merge(inhoud, annotatie, by="ID")
#fDir <-  read.table("./WCFS1_cnts.txt", header=TRUE) 
#row.names(inhoud) = inhoud[,"ID"]
```
### fName <- "WCFS1_cnts.txt"

### "WCFS1_cnts.txt"
### "WCFS1_RNA-Seq.txt"
```{r}
cnts <- read.delim(paste0(fDir), comment.char = "#") ### ,fName), comment.char="#")
```

### used for topTags identification
```{r}
row.names(cnts) <- cnts[,"ID"]
```

### Create DGEList
```{r}
#exp <- c("WCFS1.glc","WCFS1.glc","WCFS1.rib","WCFS1.rib")

group <- factor(c("WCFS1.glc.1",	"WCFS1.glc.2",	"WCFS1.rib.1"
                 , "WCFS1.rib.2", "NC8.glc.1", "NC8.glc.2"	
                 , "NC8.rib.1", "NC8.rib.2"))
y <- DGEList(counts=data[,2:9], group=group)
```
### Normalise counts
### Trimmed mean of M values : remove lowest and highest values
### (percentile) and calculate mean
```{r}
y <- calcNormFactors(y, method="TMM" )
keep <- filterByExpr(y)
x <- y[keep, , keep.lib.sizes=FALSE]
x$samples$lib.size = colSums(x$counts)
```
### Check statistics
```{r}
print("Count statistics")
print(summary(y$counts))
print(y$samples)
```
### Create design matrix
```{r}
design <- model.matrix(~0+group, data=y$samples)
colnames(design) <- levels(y$samples$group)
print(design)
```
### Estimate Dispersion

Ik weet niet zeker wat hier de goede methode is. 
y <- estimateGLMCommonDisp(y,design)
y <- estimateGLMTrendedDisp(y,design, method="power")
y <- estimateGLMTagwiseDisp(y,design)
is uit het voorbeeld.
Philip heeft genoemd dat:

#y <- estimateDisp(y, design)

beter is
```{r}
y <- estimateGLMCommonDisp(y,design)
y <- estimateGLMTrendedDisp(y,design, method="power")
y <- estimateGLMTagwiseDisp(y,design)

#y <- estimateDisp(y, design)

```
### Plot results
```{r}
pdf(paste0(fDir,"test.pdf"))
plotMDS(y)
plotBCV(y)
dev.off()
```
### Fit data
```{r}
fit <- glmFit(y,design)
```
### Determine fold changes
```{r}
mc <- makeContrasts(exp.r=WCFS1.glc-WCFS1.rib, levels=design)

fit <- glmLRT(fit, contrast=mc)
```
### Print top tags
```{r}
res<-topTags(fit)
print(res)
```
### Correlatie van de data
```{r}
cor.test(y$counts[,1], y$counts[,2])
cor.test(y$counts[,3], y$counts[,4])
```
### plot van diviance en dispersion
```{r}
plot(deviance(fit))
plot(fit$dispersion)
```
### Alle p-values
```{r}
PV <- fit[["table"]][["PValue"]]
head(PV, 100)
```
### Alle fold changes in log
```{r}
logFC <- fit[["table"]][["logFC"]]
head(logFC, 100)
```
### Alle fold changes normaal
```{r}
FC <- logFC^2
head(FC, 100)
```
### Test ding
```{r}
summary(de <- decideTestsDGE(fit))

detags <- rownames(y)[as.logical(de)]
plotSmear(fit, de.tags = detags)
abline(h=c(-1,1), col="blue")
```
